---
title: "Journal (Basics for Data Science )"
author: "Ossama Ahmed"
date: "2020-11-30"
output:
  html_document:
    toc: true
    toc_float: true
    collapsed: false
    number_sections: true
    toc_depth: 3
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(message=FALSE,warning=FALSE, cache=TRUE)
```


# Important Info

Last compiled: `r Sys.Date()`
Challange for chapter 2 is done 


# R Learning

## How to make a Histogram

```{r}
samples <- rnorm(100, mean=0, sd=1)
hist(samples)
```
## web scraping
 In general, web scraping in R (or in any other language) boils down to the following three steps:

* Get the HTML for the web page that you want to scrape
* Decide what part of the page you want to read and find out what HTML/CSS you need to select it
* Select the HTML and analyze it in the way you need

# Business Data Science Basics Challenges

## Chapter 2 Challenges

Preparing the data ...

```{r}
# 1.0 Load libraries ----
library(tidyverse)
library(readxl)
library(lubridate)
# 2.0 Importing Files ----

bikes_tbl <- read_excel(path = "DS_101/00_data/01_bike_sales/01_raw_data/bikes.xlsx")
orderlines_tbl <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/orderlines.xlsx")
bikeshops_tbl  <- read_excel("DS_101/00_data/01_bike_sales/01_raw_data/bikeshops.xlsx")

# 4.0 Joining Data ----

bike_orderlines_joined_tbl <- orderlines_tbl %>%
  left_join(bikes_tbl, by = c("product.id" = "bike.id")) %>%
  left_join(bikeshops_tbl, by = c("customer.id" = "bikeshop.id"))

# 5.0 Wrangling Data ----

bike_orderlines_wrangled_tbl <- bike_orderlines_joined_tbl %>%
  # 5.1 Separate category name
  separate(col    = category,
           into   = c("category.1", "category.2", "category.3"),
           sep    = " - ") %>%
  
  # 5.2 Add the total price (price * quantity) 
  # Add a column to a tibble that uses a formula-style calculation of other columns
  mutate(total.price = price * quantity) %>%
  
  # 5.3 Optional: Reorganize. Using select to grab or remove unnecessary columns
  # 5.3.1 by exact column name
  select(-...1, -gender) %>%
  
  # 5.3.2 by a pattern
  # You can use the select_helpers to define patterns. 
  # Type ?ends_with and click on Select helpers in the documentation
  select(-ends_with(".id")) %>%
  
  # 5.3.3 Actually we need the column "order.id". Let's bind it back to the data
  bind_cols(bike_orderlines_joined_tbl %>% select(order.id)) %>% 
  
  # 5.3.4 You can reorder the data by selecting the columns in your desired order.
  # You can use select_helpers like contains() or everything()
  select(order.id, contains("order"), contains("model"), contains("category"),
         price, quantity, total.price,
         everything()) %>%
  
  # 5.4 Rename columns because we actually wanted underscores instead of the dots
  # (one at the time vs. multiple at once)
  rename(bikeshop = name) %>%
  set_names(names(.) %>% str_replace_all("\\.", "_"))

```

### Challenge 1 
```{r}
bike_orderlines_wrangled_state_tbl <- bike_orderlines_wrangled_tbl  %>%
  # 5.1 Separate category name
  separate(col    = location,
           into   = c("city", "state"),
           sep    = ", ")
  # Analyze
sales_by_loc_tbl <- bike_orderlines_wrangled_state_tbl %>%
# Select columns
select(state, total_price) %>%
  
  # Grouping by year and summarizing sales
  group_by(state) %>% 
  summarize(sales = sum(total_price)) %>%
  
  # Add a column that turns the numbers into a currency format 
  mutate(sales_loc = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))
sales_by_loc_tbl  %>%
  
  # Setup canvas with the columns year (x-axis) and sales (y-axis)
  ggplot(aes(x = state, y = sales)) +
  
  # Geometries
  geom_col(fill = "#2DC6D6") + # Use geom_col for a bar plot
  geom_label(aes(label = sales_loc)) + # Adding labels to the bars
  geom_smooth(method = "lm", se = FALSE) + # Adding a trendline
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +

  labs(
    title    = "Revenue by State",
    subtitle = "",
    x = "", # Override defaults for x and y
    y = "Revenue"
  )
```

### Challenge 2
```{r}
sales_by_year_loc_tbl <- bike_orderlines_wrangled_state_tbl %>%
  
  # Select columns and add a year
  select(order_date, total_price, state) %>%
  mutate(year = year(order_date)) %>%
  
  # Group by and summarize year and states
  group_by(year, state) %>%
  summarise(sales = sum(total_price)) %>%
  ungroup() %>%
  
  # Format $ Text
  mutate(sales_text = scales::dollar(sales, big.mark = ".", 
                                     decimal.mark = ",", 
                                     prefix = "", 
                                     suffix = " €"))
# Step 2 - Visualize


sales_by_year_loc_tbl %>%
  
  # Set up x, y, fill
  ggplot(aes(x = year, y = sales, fill = state)) +
  
  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot
  
  # Facet
  facet_wrap(~ state) +
  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " €")) +
  
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  
  labs(
    title = "Revenue by year and State",
    subtitle = "",
    fill = "Main category" # Changes the legend name
  )
```

## Chapter 3 Challenges

### Challenge 1 
```{r}
# Libraries 
library(httr)
library(jsonlite)
library(dplyr)
library(keyring)
library("data.table")
library(tidyverse) # Main Package - Loads dplyr, purrr, etc.
library(rvest)     # HTML Hacking & Web Scraping
library(xopen)     # Quickly opening URLs
library(jsonlite)  # converts JSON files to R objects
library(glue)      # concatenate strings
library(stringi)   # character string/text processing
library(lubridate)
library(anytime)
# Securing My API 
#keyring::key_set("MYAPI")

# get data from API
my_url <- paste0("http://dataservice.accuweather.com/forecasts/",
                 "v1/daily/5day/178556?apikey=", 
                 key_get("MYAPI"))
my_raw_result <- httr::GET(my_url)
my_content <- httr::content(my_raw_result, as = 'text') %>%
  fromJSON()
forcast <- my_content$DailyForecasts

Weather <- forcast %>% select(EpochDate,Temperature)
Temperature<- Weather %>% select(Temperature)%>% as.data.table()  
date <- Weather %>% select(EpochDate) %>% as_tibble()

date <- map(date, anytime)
date<- date %>% as_tibble()
Max_Temp <- Temperature %>% select(Temperature.Maximum.Value)
Min_Temp <- Temperature %>% select(Temperature.Minimum.Value)

names(Max_Temp)[1]<-paste("Maximum Temperature")
names(Min_Temp)[1]<-paste("Minimum Temperature")

Final_Forcast <- cbind(date,Max_Temp,Min_Temp)
Final_Forcast %>% as_tibble()
Final_Forcast %>%
  
  ggplot(aes(x = EpochDate, y = `Maximum Temperature`)) +
  
  # Geometries
  geom_col() + # Run up to here to get a stacked bar plot

  
  # Formatting
  scale_y_continuous(labels = scales::dollar_format(big.mark = ".", 
                                                    decimal.mark = ",", 
                                                    prefix = "", 
                                                    suffix = " F")) +
  labs(
    title = "Forcast for next 5 Days",
    subtitle = "Maximum Temperature",
    fill = "Main category" # Changes the legend name
  )
```

### Challenge 2

```{r}
url_home          <- "https://www.rosebikes.de/fahrr%C3%A4der"

# Read in the HTML for the entire webpage
html_home         <- read_html(url_home)

bike_family_tbl <- html_home %>%
  
  html_nodes(css = ".catalog-categories-item__link") %>%
  html_text()
  # ...and extract the information of the id attribute
 # html_attr('title')

price <- html_home %>%
  
  html_nodes(css = ".catalog-categories-item__link") %>%
  html_text() %>%
  stringr::str_replace_all("[[/.]]", replacement = "") %>%
  stringr::str_extract(pattern = "[0-9 | /,]+") %>% 
  as_tibble()
# ...and extract the information of the id attribute
# html_attr('title')

model <- html_home %>%
  
  html_nodes(css = ".catalog-categories-item__link") %>%
  html_text() %>%
  stringr::str_extract("[a-z | A-Z | /-]+") %>%
  as_tibble()
#as.numeric()
# ...and extract the information of the id attribute
# html_attr('title')


names(model)[1]<-paste("Model")
names(price)[1]<-paste("Starting Price")

Data_Final <- cbind(model,price)
Data_Final
```
